---
title: "Seasonal Correlation of the Multivariate ENSO Index and Average Indian Rainfall"
author:
- Christian S. Zang, Land Surface-Atmosphere Interactions, Technical University of Munich
- Tilda M. Zang, Slumber and Sniffing Resarch Unit, Bavarian Canine University
output: html_document
bibliography: meiair.bib
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
# we do it tidily
library(tidyverse)
# load custom functions
source("R/methods.R")
```

# Introduction

To keep us busy durig the lockdown, we looked how the multivariate ENSO index (MEI) is seasonally correlated with the average Indian rainfall (AIR). Some folks said there is some correlation [@gershunov_low-frequency_2001].

# Material and Methods

We downloaded the MEI data from https://psl.noaa.gov/enso/mei/, and the AIR data from https://crudata.uea.ac.uk/cru/data/hrg/cru_ts_4.01/crucy.1709191757.v4.01/countries/pre/. The monthly resolution of the AIR data was aggregated into two-monthly seasons to match the temporal aggregation of MEI. For each season, we computed Kendall's tau for MEI and AIR for the common time span.

```{r two-month seasons}
tms <- c("DJ", "JF", "FM", "MA", "AM", "MJ", "JJ",
         "JA", "AS", "SO", "ON", "ND")

```

```{r read enso data}
mei <- read_lines("data/meiv2.data.txt", skip = 1) %>% 
  .[-43:-46] %>% 
  read_table(col_names = c("year", tms),
             na = "-999")
```

```{r read indian precip data}
air <- read_table("data/crucy.v4.01.1901.2016.India.pre.per",
                  skip = 3, na = "-999.0")
```

```{r seasonalise air}
air <- air %>% 
  select(YEAR:DEC) %>% 
  mutate(PDEC = lag(DEC)) %>% 
  select(YEAR, PDEC, everything()) %>% 
  season_fun(tms)
```

```{r tidy data}
airmei <- air %>% gather(season, precip, -year) %>% 
  inner_join(mei %>% gather(season, mei, -year))
```

```{r seasonal correlations}
seas_corr <- airmei %>% 
  mutate(season = ordered(season, levels = tms)) %>% 
  group_by(season) %>% 
  nest() %>% 
  mutate(correlation = purrr::map_dbl(data, cor_fun))
```

# Results

We found that the correlation between MEI and AIR switches signs, with positive correlations from ON to MA, and negative correlations between MJ and SO (Figure 1).

```{r, fig.cap="Seasonal correlation (Kendall's tau) between bi-monthly mean MEI and AIR"}
seas_corr %>% 
  ggplot(aes(x = season, y = correlation)) +
  geom_col()
```


# Discussion

Well, that was exactly what we expected. I mean, look at @gershunov_low-frequency_2001!

# Autor contribution statement

CSZ conceived the study, carried out all analyses and wrote the manuscript. TMZ slept most of the time, and occassionally interrupted CSZ by licking the keyboard.

# References